Since we are working with larger object sizes, it's best to start with a fresh session. All variable names will be different from the lessons. Load the libraries (it's okay if they're already loaded) and a Seurat object with alternative clustering using the following code:
<!-- add copy of alt clustering to share with learners -->

```{r, Day3_startup, eval = FALSE}
# =========================================================================
# Independent Exercise - Day 3 Startup
# =========================================================================

# After restarting our R session, load the required libraries
library(Seurat)
library(BPCells)
library(tidyverse)

# Load in seurat object with alternative clustering results from yesterday's exercises
exso3 = readRDS('results/rdata/geo_so_sct_clustered_alt2.rds')
exso3 # check that object loaded

## NOTE - BEFORE STOPPING WORK ON THE EXERCISES REMEMBER TO POWER DOWN AND RESTART R SESSION !!!!
```

<br/>


### Day 3 Exercise 1 - examine marker genes and cell type predictions for alternative clustering

Now that we've covered how to generate emperical marker genes and use a prediction tool to identify possible cell-type labels, let's start by extending the exercise from yesterday to check what our results would look like if we had proceed with fewer or more PCs.

```{r, eval=FALSE}
## ----------------------------------------------------------
# Check what identities are set
Idents(exso3) %>% head()

res = 6 ## use same res values as previous exercise

## Note - in the example code, we use the results from "fewer" but can swap to "more" instead
exso3 = SetIdent(exso3, value = paste0('int.sct.rpca.clusters', res))
Idents(exso3) %>% head()
```



```{r, eval=FALSE}
## Generate cluster markers to see how that changes with new parameters 
exso3 = PrepSCTFindMarkers(exso3, assay = "SCT") # NOTE - this step will take some time to run
geo2_markers = FindAllMarkers(exso3, only.pos = TRUE)

# Create table of top 5 markers per cluster (using default ranking)
top_5 = geo2_markers %>% filter(p_val_adj < 0.01) %>% group_by(cluster) %>% slice_head(n = 5)
head(top_5, n = 10) # look at results
write_csv(top_5, file = paste0('results/tables/top5_marker_genes_fewer.csv'))
```
<br/>

##### Check-in Question - clustering impact on marker genes
After you generate markers for the "fewer" `pcs` option, how do the results differ from the markers found for `pcs=10` in the workshop session? What do you think would happen if you tried this with the "more" results?

<br/>

```{r, eval=FALSE}
## ----------------------------------------------------
## Next - run scCATCH predictions for alternative clustering results
geo2_catch = createscCATCH(data = exso3@assays$SCT@counts, cluster = as.character(Idents(exso3)))
catch_markers = geo2_markers %>% rename('logfc' = 'avg_log2FC')
geo2_catch@markergene = geo2_markers
geo2_catch@marker = cellmatch[cellmatch$species == 'Mouse' & cellmatch$tissue %in% c('Blood', 'Peripheral Blood', 'Muscle', 'Skeletal muscle', 'Epidermis', 'Skin'), ]
geo2_catch = findcelltype(geo2_catch)

# Check predictions
geo2_catch@celltype %>% select(cluster, cell_type, celltype_score)
```

```{r, eval=FALSE}
## ------------------------------------------------------
## Use predictions to label clusters and UMAP plot

catch_celltypes = geo2_catch@celltype %>% select(cluster, cell_type)
colnames(catch_celltypes)[2] = paste0('cell_type.',pcs,'PC.',res,'res')
new_metadata = exso3@meta.data %>% 
  left_join(catch_celltypes, 
            by = c('seurat_clusters' = 'cluster')) # using `seurat_clusters`, which will store the most recently generated cluster labels for each cell
rownames(new_metadata) = rownames(exso3@meta.data) #  We are implicitly relying on the same row order!
exso3@meta.data = new_metadata # Replace the meta.data
head(exso3@meta.data)

# Create UMAP plot with new cluster labels
catch_umap_plot = DimPlot(exso3, group.by = paste0('cell_type.',pcs,'PC.',res,'res'), 
                          label = TRUE, reduction = paste0('umap.integrated.sct.rpca_',pcs,'PC'))
catch_umap_plot
#### Question: How did the number of pcs and/or resolution change the predictions? Do you think the predictions correspond better or worse to the cluster structure we see in the UMAP?

# Save the plot to file
# output to file, including the number of PCs and resolution used to generate results
ggsave(filename = paste0('results/figures/umap_int_catch-labeled_', 
                         pcs,'PC.',res,'res','.png'),
       plot = catch_umap_plot, 
       width = 8, height = 6, units = 'in')

## Clean up session 
rm(list=names(which(unlist(eapply(.GlobalEnv, is.ggplot))))); 
rm(catch_celltypes, catch_markers, geo2_catch, geo2_markers, new_metadata, top_5); 
gc()

## (Optional) - Save copy of exso3
saveRDS(exso3, file = paste0('results/rdata/exso3_sct_integrated_alt.rds'))

## BEFORE PROCEEDING TO THE NEXT SECTION or closing window - POWER DOWN AND RESTART R SESSION

```
<br/>

### Day 3 Challenge - Run additional DE comparisons using session results

```{r, eval=FALSE}
## BEFORE PROCEEDING, MAKE SURE SESSION HAS BEEN restarted and environment is clear

# After restarting our R session, load the required libraries
library(Seurat)
library(BPCells)
library(tidyverse)

# load a copy of the final Seurat object (from end of Day 3)
exso3 = readRDS('/home/workshop/damki/ISC_R/results/rdata/exso3_sct_integrated_final.rds')

# check what identities are set for the loaded Seurat object
Idents(exso3) # expect to see Day + Cluster labels; if different identities are set, how would you change them?
## add code to change identities if needed
```

Once the cluster identities are set to the combined Day + Cluster labels, generate additional DE comparisons using the provided "psuedocode" which are comments that outline what steps are expected/needed to guide adding code to complete those outlined steps.
```{r, eval = FALSE}
## ---------------------------------------------------
## Further extension - how might you generate DE comparisons between D21 and D7 for all annotated clusters?

# check the unique cluster labels - how many clusters are present?
unique(exso3$cell_type)

## Below is comments to help guide you though the process of looping through all the comparisons of interest, for all labeled clusters
## This section from the HBC materials could be a helpful reference to aid in filling in the code: https://hbctraining.github.io/scRNA-seq_online/lessons/pseudobulk_DESeq2_scrnaseq.html

# Create list(s) of pair-wise comparisons of interest between conditions, ideally specifying Case and Control 
# (e.g. Day 7 vs Day 0, Day 21 vs Day 7, Day 21 vs Day 0)

# Use list of comparisons and cluster identities to loop through comparisons of interest (for standard DE comparisons) and save results to files
# Try to use an outer `for` loop step though the comparisons listed in previous step and an inner `for` loop to cycle through all the clusters

# Bonus - summarize the DE results for the comparison, choosing a reasonable set of thresholds to call the number of DE genes, & save summary to file

# Clean up session. Caution - this next command clears your environment of all objects
rm(list = ls())
gc()

## BEFORE CLOSING WINDOW - POWER DOWN AND RESTART R SESSION !!!!

```

</br>
